
---
title: "Homework 1"
author: "Group M: Billo, Pizzignacco, El Gataa, Shahazad"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Load necessary libraries here if any
#### TO USE GITHUB ####
install.packages("usethis")
library(usethis)

#### other libraries ####
library(dplyr)
```

## Introduction

*Briefly describe the purpose of the homework or tasks to be accomplished.*

## Block B

### CS: Chapter 1, exercises 1.1, 1.6


```{r}
# Placeholder for code for Exercise B1
```
### CS: Chapter 3, exercises 3.5, 3.6


```{r}
# Placeholder for code for Exercise B1
```

### FSDS: Chapter 2, exercises 2.8, 2.16, 2.21, 2.26, 2.52, 2.53, 2.70


```{r exercise_b2}
# Placeholder for code for Exercise B2
```

### FSDS: Chapter 3, exercises 3.18, 3.28, 3.24 


```{r exercise_b3}
# Placeholder for code for Exercise B3
```

### FSDS: Chapter 4, exercises 4.14, 4.16, 4.48 

# Exercise 4.14
```{r 4.14}
data = read.table("https://stat4ds.rwth-aachen.de/data/Students.dat", header = TRUE)
summary(data)
#a
x_bar = mean(data$tv)
s = sd(data$tv)
# H0: the mean is 7.2
# H1: the mean is different from 7.2
z = qnorm(0.975) # having n > 60, we can assume normality
SE = s/sqrt(length(data$tv))
CI = x_bar + c(-1, 1)*z*SE
# we can say that 95% of the students on average spend between 5.30 and almost 9 hours watching TV per week

#b
only_male = data %>% filter(gender == 0)
only_female = data %>% filter(gender == 1)

boxplot(only_male$tv, only_female$tv, names = c("Males", "Females"), ylab = "weekly hours watching tv")
```
assuming
-independent populations
-equal variances
-normality -> weakest assumption, barely 30 observations
--> use t student's distribution
```{r}
x_bar_male = mean(only_male$tv)
x_bar_female = mean(only_female$tv)
s_male = sd(only_male$tv)
s_female = sd(only_female$tv)
n_male = length(only_male$tv)
n_female = length(only_female$tv)

# assuming equal variance
s_p = ((n_male -1)*(s_male^2) + (n_female - 1)*(s_female^2))/(n_male + n_female -1)
# t = (x_bar_male - x_bar_female)/(s_p*(1/n_male + 1/n_female))
SE_diff = (x_bar_male - x_bar_female)/((s_p^2)*(1/n_male + 1/n_female))
mean_diff = x_bar_male - x_bar_female
z = qt(0.975, n_male + n_female - 2)
CI_diff = mean_diff + c(-1, 1)*z*SE_diff

# Test for equality of means
t.test(only_male$tv, only_female$tv, conf.level = 0.05)
```

We can not state that the variances are significantly different betweent the 2 groups
However the confidence interval seems to suggest that females watch slightly more tv than males.

```{r}
# assuming unequal variance
SE_diff = sqrt((s_male^2)/n_male + (s_female^2)/n_female)
# t = (x_bar_male - x_bar_female)/(s_p*(1/n_male + 1/n_female))
mean_diff = x_bar_male - x_bar_female
z = qt(0.975, n_male + n_female - 2)
CI_diff_uvars = mean_diff + c(-1, 1)*z*SE_diff
```

Now we the interval crosses 0?? how is such a different result possible?

# Exercise 4.16
```{r 4.16}

data = read.table("https://stat4ds.rwth-aachen.de/data/Substance.dat", header = TRUE)

# compare alcohol users and non-users
alpha = 0.05
n = sum(data$count)
# find the total number of students that have or haven't used alcohol
N_alcohol_total = sum(data[data$alcohol == "yes", 4])
N_NOalcohol_total = sum(data[data$alcohol == "no", 4])
N_marijuana = sum(data[data$alcohol == "yes" & data$marijuana == "yes", 4])
N_NOalcohol_marijuana = sum(data[data$alcohol == "no" & data$marijuana == "yes", 4])

pi_1_hat = N_marijuana/N_alcohol_total
pi_2_hat = N_NOalcohol_marijuana/N_NOalcohol_total

# by hand
z = qnorm(1-alpha/2)
SE = (pi_1_hat*(1 - pi_1_hat)/N_alcohol_total + pi_2_hat*(1 - pi_2_hat)/N_NOalcohol_total)
CI_prop = (pi_1_hat - pi_2_hat) + c(-1, 1)*z*sqrt(SE)

#with software
prop.test(c(955, 5), c(1949, 327), conf.level = alpha, correct = F)
```
Interpretation: there seems to be a significant mean difference in the use of marijuana 
between students that used alcohol before and students who didn't.


# Exercise 4.48

## Given

\[
\text{Given } \quad SE = \sqrt{\frac{\hat{p}(1 - \hat{p})}{n}}
\]

With 95% likelihood, find:
$$ 
SE \leq \frac{1}{\sqrt{n}} \quad \text{when } \hat{p} = 0.5 
$$
This is where \( SE \) is maximized.

\[
SE = \sqrt{\frac{0.5 \times (0.5)}{n}} = \frac{\sqrt{0.25}}{\sqrt{n}} = \frac{0.5}{\sqrt{n}} = \frac{1}{2 \sqrt{n}}
\]

## Example: When $\hat{p} \neq 0.5$

\[
SE = \sqrt{\frac{0.3 \times (1 - 0.3)}{n}} = \sqrt{\frac{0.21}{n}} = \frac{0.458}{\sqrt{n}} \approx \frac{0.91}{2\sqrt{n}}
\]

## Example: When $\hat{p} = 0.2$

\[
SE = \sqrt{\frac{0.2 \times (0.8)}{n}} = \sqrt{\frac{0.16}{n}} = \frac{0.4}{\sqrt{n}} = \frac{0.8}{2\sqrt{n}}
\]

## Example: When $\hat{p} = 0.7$

\[
SE = \sqrt{\frac{0.7 \times (0.3)}{n}} = \sqrt{\frac{0.21}{n}} = \frac{0.458}{\sqrt{n}} \approx \frac{0.91}{2\sqrt{n}}
\]

## Observation

\[
\text{Note that when both } \hat{p} < 0.5 \text{ and } \hat{p} > 0.5 \text{ the numerator decreases, along with the standard error.}
\]

## For Maximum Standard Error

Set the maximum \( SE \) to be within margin \( M \).

\[
\frac{1}{\sqrt{n}} = M
\]

\[
\Rightarrow \frac{1}{M} = \sqrt{n} \Rightarrow n = \frac{1}{M^2}
\]

As long as \( n \geq \frac{1}{M^2} \), our error will be within \( M \).



### FSDS: Chapter 5, exercises 5.2, 5.12, 5.50 ###

```{r exercise_b5}
# Placeholder for code for Exercise B5
```
