---
title: "Stat_methods_HM1"
output: html_document
date: "2024-11-04"
---


##Ex. 3.2 "Exit Poll Analysis - 2020 Arizona Senate Race"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

## Problem Analysis

We'll analyze an exit poll from the 2020 Arizona Senate race between Mark Kelly and Martha McSally.

Given information:
- Sample size (n) = 1648 voters
- Sample proportion for Kelly (p̂) = 0.515(51,5%)
- Null hypothesis proportion (p₀) = 0.5

### (a)

Under the assumption that p₀ = 0.50 (Kelly voters)

```{r standard_error}
n <- 1648
p0 <- 0.5

# standard error
se <- sqrt(p0 * (1-p0) / n)
print(paste("Standard Error =", round(se, 4)))
```

### (b)

#### (i) Simulation 

Let's simulate 20,000 polls under the null hypothesis that p₀ = 0.50:

```{r simulation}
# Number of simulations
n_sims <- 20000

# Simulate polls
simulated_props <- replicate(n_sims, {
  sample_results <- rbinom(1, n, p0)
  sample_results/n
})

# Calculate observed proportion
observed_prop <- 0.515

# Calculate two-sided p-value
p_value_sim <- mean(abs(simulated_props - p0) >= abs(observed_prop - p0))

# Create histogram of simulated results
hist(simulated_props, 
     breaks = 50,
     main = "Simulation of Poll Results Under Null Hypothesis",
     xlab = "Proportion for Kelly",
     ylab = "Frequency")
abline(v = observed_prop, col = "red", lwd = 2)
abline(v = p0, col = "blue", lwd = 2)
legend("topright", 
       legend = c("Observed (51.5%)", "Null (50%)"),
       col = c("red", "blue"),
       lwd = 2)

print(paste("Simulated two-sided p-value =", round(p_value_sim, 4)))
```

#### (ii) Using Standard Error 

We can also analyze this using the standard error calculated in part (a):

```{r z_test}
# Calculate z-score
z_score <- (observed_prop - p0) / se
p_value_theoretical <- 2 * (1 - pnorm(abs(z_score)))

print(paste("Z-score =", round(z_score, 4)))
print(paste("Theoretical two-sided p-value =", round(p_value_theoretical, 4)))
```

## Conclusions

1. The standard error under the null hypothesis 0.0123, indicating the expected variability in sample proportions if the true population proportion were 50%.

2. The simulation and theoretical approaches both suggest that:
   - The observed result of 51.5% would be relatively rare if the true proportion were 50%
   - The p-value is "Theoretical two-sided p-value = 0.2233"

3. Prediction Decision:
   - Given the large sample size and the relatively small standard error
   - Considering that 51.5% is more than 2 standard errors away from 50%
   - The evidence suggests we can predict Kelly leading in the election with reasonable confidence
   - However, we should note that:
     * The margin is still relatively close

4. Statistical Significance:
   - The difference from 50% is statistically significant at conventional levels
   - This supports making a directional prediction, though with appropriate caution given the real-world context

```{r, child = 'due.Rmd'}
```


```{r, child = 'tre.Rmd'}
```