---
title: "terzo"
output: html_document
date: "2024-11-04"
---

### Terzo esercizio (3.24)

```{r}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

## Possiblee Pop. Distribution

Distribution for the number of alcoholic drinks consumed in the past day (Y) right-skewed distribution since:
- Many people don't drink (0 drinks)
- Some people have a few drinks
- A smaller number of people have many drinks
- Negative values are impossible

```{r population}
# Create population distribution parameters
n_population <- 100000  # Large population size

prob_zero <- 0.4  # 40% of people don't drink

shape <- 1.5
rate <- 1

# Generate population data
zeros <- rep(0, prob_zero * n_population)
drinkers <- rgamma(n = (1 - prob_zero) * n_population, shape = shape, rate = rate)
population <- c(zeros, drinkers)

# Calculate population parameters (mean and standard dev)
pop_mean <- mean(population)
pop_sd <- sd(population)

# Print population parameters
print(paste("Population Mean:", round(pop_mean, 3)))
print(paste("Population SD:", round(pop_sd, 3)))

# Visualize population distribution
hist(population, 
     breaks = 50,
     main = "Population Distribution of Daily Drinks",
     xlab = "Number of Drinks",
     ylab = "Frequency",
     xlim = c(0, 10))
```

## A. Single Sample Analysis

Let's draw one random sample of size 1000 from this population:

```{r single_sample}
# single sample
sample_size <- 1000
single_sample <- sample(population, size = sample_size, replace = TRUE)

# sample statistics
sample_mean <- mean(single_sample)
sample_sd <- sd(single_sample)

# comparisons
print("Sample vs Population Statistics:")
print(paste("Sample Mean:", round(sample_mean, 3), "vs Population Mean:", round(pop_mean, 3)))
print(paste("Sample SD:", round(sample_sd, 3), "vs Population SD:", round(pop_sd, 3)))

# Visualize sample distribution
hist(single_sample, 
     breaks = 30,
     main = "Sample Distribution (n=1000)",
     xlab = "Number of Drinks",
     ylab = "Frequency",
     xlim = c(0, 10))
```

## B. Sampling Distribution Analysis

simulate 10,000 samples of size 1000 each to approximate the sampling distribution:

```{r sampling_distribution}
# Simulate 10000 samples
n_simulations <- 10000
sample_means <- numeric(n_simulations)

for(i in 1:n_simulations) {
    sample_means[i] <- mean(sample(population, size = sample_size, replace = TRUE))
}

# sampling distribution statistics
sampling_mean <- mean(sample_means)
sampling_sd <- sd(sample_means)

# theoretical standard error
theoretical_se <- pop_sd/sqrt(sample_size)

# Print results
print("Sampling Distribution Statistics:")
print(paste("Simulated Mean of Sample Means:", round(sampling_mean, 4)))
print(paste("Theoretical Mean (Population Mean):", round(pop_mean, 4)))
print(paste("Simulated Standard Error:", round(sampling_sd, 4)))
print(paste("Theoretical Standard Error:", round(theoretical_se, 4)))

# Visualize sampling distribution
hist(sample_means, 
     breaks = 50,
     main = "Sampling Distribution of Sample Mean",
     xlab = "Sample Mean Number of Drinks",
     ylab = "Frequency",
     probability = TRUE)
curve(dnorm(x, mean = pop_mean, sd = theoretical_se), 
      add = TRUE, 
      col = "red", 
      lwd = 2)
legend("topright", 
       legend = c("Simulated Distribution", "Normal Approximation"),
       col = c("grey", "red"),
       lwd = c(1, 2))
```

## Key Observations

1. Population Distribution:
   - Right-skewed distribution
   - 40% of people have 0 drinks
   - Mean and standard deviation 
   - Represents the true distribution of daily drinks in the population

2. Single Sample (n = 1000):
   - Shows similar shape to population
   - Sample mean and standard deviation are close to population values
   - Demonstrates what we might see in a real survey (real data)

3. Sampling Distribution:
   - Nearly normal in shape (Central Limit Theorem)
   - Centered very close to the population mean
   - Standard error (SD of sampling distribution) is close to theoretical value
   - Represents the distribution of sample means if we repeated the survey many times
   - Much narrower than the population distribution, showing that sample means vary less than individual observations

4. Theory vs Simulation:
   - Simulated results closely match theoretical expectations
   - Small differences due to random variation in simulation
   - Confirms the accuracy of our theoretical understanding of sampling distributions

so we can demostrat that the sampling distribution of the mean becomes approximately normal (due to the CLT) even though the underlying population distribution is highly skewed.











